---
title: "ps_4"
author: "Arnav Srivastava"
date: "3/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# as per basic usage of tidycensus link, importing relavant libraries for
# tidycensus and using personalized census key to explore census data. 
# Also upload gt for gt table for 5c)

library(tidycensus)
library(tidyverse)
library(gt)

# applies proper argument to census_api_key to run census key multiple times and
# activate (without restarting R... need to do this to avoid warning)

census_api_key("17e6fac7c29f16aa547f2cdb01baeb61a82db36a", install = TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")



# this code line prevents R from printing numbers in standard notation, for
# cleaner/desired numeric display formatting for madlibs and question 2 mappin

options("scipen"=100, "digits"=4)



# use get_acs to get median household income at the country level for 2018 for
# madlibs

income <- get_acs(geography = "county", 
              variables = c(medincome = "B19013_001"),  
              year = 2018)



# draw acs data to map median income by county in California for Question 2

calif <- get_acs(geography = "county",
                 state = "California",
                 variables = c(medincome = "B19013_001"),  
                 year = 2018,
                 geometry = TRUE)


```

## Mad Libs

```{r madlibs, echo = FALSE}

# q1; copying function from mad lib 1
# removed +1 after na.rm assignment (illogical)
# put a closed parenthesis for the return arg
# added +1 to return since want sd +1 of a numeric var

sd_plus_one <- function(var, na.rm = TRUE){
  x <- sd(var, na.rm = na.rm)
  return(x + 1)
  }

q1 <- sd_plus_one(income$estimate)


# q2; copying function from mad lib 2
# removed %>%, and set list x to contain 2 parts of countystate
# then, had pluck take the 2nd item of list, making return irrelavant


get_state <- function(countystate){
  x <- str_split_fixed(countystate, ", ", 2) 
  pluck(x, 2)
 }

q2 <- get_state("Treasure County, Montana")



# q3; copying function from mad lib 3 
# change entry arg of upper_bound to geoid,
# so that actually used in function, and also 
# removed extra parenthesis after pull() function

upper_bound <- function(geoid){
 income %>% 
    filter(GEOID == geoid) %>% 
    mutate(estimate + moe) %>% 
    pull()
}

q3 <- upper_bound(geoid = "01055")
```

**ML 1)** The standard deviation of median income, plus one, is `r q1`.

**ML 2)** Treasure County, Montana is located in the state of `r q2`.

**ML 3)** The upper bound for median income in Etowah County, Alabama (geoid = “01055”) is `r q3`.


## Question 2: Mapping

```{r california_map, echo = FALSE}

# make ggplot filling by median income estimate, and use scale_fill to get more
# distinguishable colors. use geom_sf to get the correct geometry of the map.
# Finally, used labs to label graph title and legend

calif %>%
  ggplot(aes(fill = estimate)) +
  geom_sf() +
  scale_fill_viridis_c(option = "plasma",
                       direction = -1)  +
  labs(title = "Median Income by County in California",
       caption = "Source: American Communities Survey 2018",
       fill = "Median Income") +
  theme_void()


```


## Question 3: Publishing to Rpubs

Here is the link to my RPubs Graphic: https://rpubs.com/arnavinator/580919


## Question 4: Understanding Tufte Reading


## Question 5: Writing a Function

### 5A
```{r 5a, echo = FALSE}

# create get_letters func which samples from letters n times, and then returns the
# minimum (since R recognizes alphabetical order), essentially returning the
# earliest alphabet. Made sure to set replace = TRUE so that we can roll the
# dice more than 26 times and each letter is a possiblity every roll like an
# actual die

get_letters <- function(n) {
 min(sample(letters, 
            size = n, 
            replace = TRUE)
     )
}


# calls get_letters with n = 3 in the RMD file

```



Call to function with n = 3:
`r get_letters(3)`


### 5B
```{r 5b, echo = FALSE}

# creates roll_alpha_dice, where argument n is set for the number of people who
# are rolling dice (number of times the map_char function runs get_letters).
# Meanwhile, the arg letter_count = 3 since we want the function to give three
# letters per person

roll_alpha_dice <- function(n, letter_count = 3) {
  map_chr(1:n, ~get_letters(letter_count))
}

```

(no output requested)

### 5c
```{r 5c, echo = FALSE}

# made a tibble where the dice was rolled for each TF to give their "lowest"
# letter, using rep to make alpha_dice for 4 people 100 times

tf_game <- tibble(game = 1:100,
                  results = map(rep(4, 100), roll_alpha_dice))


# make a gt table slicing the alpha dice game to the first five turns and
# formatting into a gt table.

tf_game %>% 
  slice(1:5) %>% 
  gt(auto_align = TRUE) %>% 
  tab_header(title = "Alphabet Dice Game") %>% 
  cols_label(results = "Alyssa, Kaneesha, Mitchell and June")



# mutated tf_game to have a vowel column, which uses map_chr to extract all of
# Alyssa's rolls, and then assigns TRUE if Alyssa's roll in that round was
# vowel, else applies FALSE to the vowel column

tf_game <- tf_game %>% 
  mutate(vowel = map_lgl(map_chr(tf_game$results, 1), 
                         ~ ifelse(.[[1]] %in% c("a", "e", "i", "o", "u"), 
                                  TRUE, 
                                  FALSE)
                         )
         )



# R is able to treat TRUE as 1 and FALSE as 0, so simply summing vowel column
# allows us count the number of times Alyssa had TRUE, ie had vowels, in the
# game

sum <- tf_game %>% 
  summarize(sum = sum(vowel)) %>% 
  pull(1)

```
 
 

Alyssa had `r sum` vowels in her game of Alpha Dice.

### 5D

```{r 5d, echo = FALSE}

# creates a table which tests 10000 rolls, and creates a column which counts
# often all of them could roll a vowel in one round of Alpha Dice (with the
# assignment mechanism similar to the vowel column)

surprised <- tibble(game = 1:10000,
                    results = map(rep(4, 10000), roll_alpha_dice))
surprised <- surprised %>% 
  mutate(perfect_vowel = map_lgl(surprised$results, 
                         ~ ifelse(all(. %in% c("a", "e", "i", "o", "u")), 
                                  TRUE,
                                  FALSE
                                  )
                         )
         )

# similar to how Alyssa's vowel was counted, we count the total number of times
# all 4 TFs had rolled a vowel and calculated the mean to determine the
# likelihood of this event - determining how likely this event is to occur!

probab <- surprised %>%
  pull(perfect_vowel) %>%
  mean()

```


Probability that they all roll a vowel: `r probab`


### 5E

```{r 5e, echo = FALSE}

dist <- tibble(results = map_chr(rep(1, 1000), roll_alpha_dice)) %>% 
  count(results)

dist

ggplot(dist, aes(x = results, y = n)) +
  geom_col() +
  labs(title = "Alpha Dice Letter Distribution",
       x = "Letter",
       y = "Letter Count")
```

Collaborators: None.
