---
title: "ps_4"
author: "Arnav Srivastava"
date: "3/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# as per basic usage of tidycensus link, importing relavant libraries for
# tidycensus and using personalized census key to explore census data. 
# Also upload gt for gt table for 5c)

library(tidycensus)
library(tidyverse)
library(gt)

# applies proper argument to census_api_key to run census key multiple times and
# activate (without restarting R... need to do this to avoid warning)

census_api_key("17e6fac7c29f16aa547f2cdb01baeb61a82db36a", install = TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")



# this code line prevents R from printing numbers in standard notation, for
# cleaner/desired numeric display formatting for madlibs and question 2 mappin

options("scipen"=100, "digits"=4)



# use get_acs to get median household income at the country level for 2018 for
# madlibs

income <- get_acs(geography = "county", 
              variables = c(medincome = "B19013_001"),  
              year = 2018)



# draw acs data to map median income by county in California for Question 2

calif <- get_acs(geography = "county",
                 state = "California",
                 variables = c(medincome = "B19013_001"),  
                 year = 2018,
                 geometry = TRUE)


```

## Mad Libs

```{r madlibs, echo = FALSE}

# q1; copying function from mad lib 1
# removed +1 after na.rm assignment (illogical)
# put a closed parenthesis for the return arg
# added +1 to return since want sd +1 of a numeric var

sd_plus_one <- function(var, na.rm = TRUE){
  x <- sd(var, na.rm = na.rm)
  return(x + 1)
  }

q1 <- sd_plus_one(income$estimate)


# q2; copying function from mad lib 2
# removed %>%, and set list x to contain 2 parts of countystate
# then, had pluck take the 2nd item of list, making return irrelavant


get_state <- function(countystate){
  x <- str_split_fixed(countystate, ", ", 2) 
  pluck(x, 2)
 }

q2 <- get_state("Treasure County, Montana")



# q3; copying function from mad lib 3 
# change entry arg of upper_bound to geoid,
# so that actually used in function, and also 
# removed extra parenthesis after pull() function

upper_bound <- function(geoid){
 income %>% 
    filter(GEOID == geoid) %>% 
    mutate(estimate + moe) %>% 
    pull()
}

q3 <- upper_bound(geoid = "01055")
```

**ML 1)** The standard deviation of median income, plus one, is `r q1`.

**ML 2)** Treasure County, Montana is located in the state of `r q2`.

**ML 3)** The upper bound for median income in Etowah County, Alabama (geoid = “01055”) is `r q3`.


## Question 2: Mapping

```{r california_map, echo = FALSE}

# make ggplot filling by median income estimate, and use scale_fill to get more
# distinguishable colors. use geom_sf to get the correct geometry of the map.
# Finally, used labs to label graph title and legend

calif %>%
  ggplot(aes(fill = estimate)) +
  geom_sf() +
  scale_fill_viridis_c(option = "plasma",
                       direction = -1)  +
  labs(title = "Median Income by County in California",
       caption = "Source: American Communities Survey 2018",
       fill = "Median Income") +
  theme_void()


```


## Question 3: Publishing to Rpubs

Here is the link to my RPubs Graphic: https://rpubs.com/arnavinator/580919


## Question 4: Understanding Tufte Reading

There are a few aspects this PowerPoint could improve on. 
The graphic on slide 2 does not contain numeric detail, as the graphic is more space taking compared to simply listing California's ranking in how high it's state/local taxes are nationwide and compared to the US (where precise numbers can be provided instead of arbitrarily comparing higher and lower states with non-labelled bar graphs).
Additionally, in slide 14, the pie graph on income tax base broadening is quite uninformative: it is difficult to tell relative differences between inroads on income tax, and this information could be better portrayed in a numeric table or flowchart.
In slide 5, explaining criteria of tax systems is limited by bullet points, as the audience cannot understand what assumptions regarding a good economy are being made, and there is little information on how "impact of economic activity" is defined. 
Meanwhile, with slide 7, exploring the negatives of the tax system are restrained by the bullet points, as one is not able to understand why volatility or marginal rates matter as cons of the California tax-system. Bullet points, in an attempt to condense information, make it difficult to understand conclusions in California's tax reform.
In order to better transmit this PowerPoint into a handout, graphs should have key data ideas be more tangible by including exact numbers and describing trends. Furthermore, bullet points in explaining tax system criteria require greater detail on core assumptions to allow for a better understanding of the criteria and conclusions.

## Question 5: Writing a Function

### 5A
```{r 5a, echo = FALSE}

# create get_letters func which samples from letters n times, and then returns the
# minimum (since R recognizes alphabetical order), essentially returning the
# earliest alphabet. Made sure to set replace = TRUE so that we can roll the
# dice more than 26 times and each letter is a possiblity every roll like an
# actual die

get_letters <- function(n) {
 min(sample(letters, 
            size = n, 
            replace = TRUE)
     )
}


# calls get_letters with n = 3 in the RMD file

```



Call to get_letters with n = 3 yields:
`r get_letters(3)`


### 5B
```{r 5b, echo = FALSE}

# creates roll_alpha_dice, where argument n is set for the number of people who
# are rolling dice (number of times the map_char function runs get_letters).
# Meanwhile, the arg letter_count = 3 since we want the function to give three
# letters per person

roll_alpha_dice <- function(n, letter_count = 3) {
  map_chr(1:n, ~get_letters(letter_count))
}

```

(no output requested)

### 5C
```{r 5c, echo = FALSE}

# made a tibble where the dice was rolled for each TF to give their "lowest"
# letter, using rep to make alpha_dice for 4 people 100 times

tf_game <- tibble(game = 1:100,
                  results = map(rep(4, 100), roll_alpha_dice))


# make a gt table slicing the alpha dice game to the first five turns and
# formatting into a gt table.

tf_game %>% 
  slice(1:5) %>% 
  gt(auto_align = TRUE) %>% 
  tab_header(title = "Alphabet Dice Game") %>% 
  cols_label(results = "Alyssa, Kaneesha, Mitchell and June")



# mutated tf_game to have a vowel column, which uses map_chr to extract all of
# Alyssa's rolls, and then assigns TRUE if Alyssa's roll in that round was
# vowel, else applies FALSE to the vowel column

tf_game <- tf_game %>% 
  mutate(vowel = map_lgl(map_chr(tf_game$results, 1), 
                         ~ ifelse(.[[1]] %in% c("a", "e", "i", "o", "u"), 
                                  TRUE, 
                                  FALSE)
                         )
         )



# R is able to treat TRUE as 1 and FALSE as 0, so simply summing vowel column
# allows us count the number of times Alyssa had TRUE, ie had vowels, in the
# game

sum <- tf_game %>% 
  summarize(sum = sum(vowel)) %>% 
  pull(1)

```
 
 

Alyssa had `r sum` vowels in her game of Alpha Dice.

### 5D

```{r 5d, echo = FALSE}

# creates a table which tests 10000 rolls, and creates a column which counts
# often all of them could roll a vowel in one round of Alpha Dice (with the
# assignment mechanism similar to the vowel column)

surprised <- tibble(game = 1:10000,
                    results = map(rep(4, 10000), roll_alpha_dice))
surprised <- surprised %>% 
  mutate(perfect_vowel = map_lgl(surprised$results, 
                         ~ ifelse(all(. %in% c("a", "e", "i", "o", "u")), 
                                  TRUE,
                                  FALSE
                                  )
                         )
         )

# similar to how Alyssa's vowel was counted, we count the total number of times
# all 4 TFs had rolled a vowel and calculated the mean to determine the
# likelihood of this event - determining how likely this event is to occur!

probab <- surprised %>%
  pull(perfect_vowel) %>%
  mean()

```


Probability that all TFs roll a vowel: `r probab`


### 5E

```{r 5e, echo = FALSE}

# creates a table using map_chr for 1 tf, giving a col with the letter winner
# each round, and then counts the number of letters for each round

dist <- tibble(results = map_chr(rep(1, 1000), roll_alpha_dice)) %>% 
  count(results)


# plots letter rolled vs number of times the letter was rolled

ggplot(dist, aes(x = results, y = n)) +
  geom_col() +
  labs(title = "3 Roll Alpha Dice Letter Distribution",
       x = "Letter",
       y = "Letter Count")
```

### 5F
```{r 5f, echo = FALSE}

# since we want to check for distributional change for rolling 5 dice, I copied
# over roll_alpha_dice and modified it accordingly, changing letter_count to 5

roll_alpha_dice_5 <- function(n, letter_count = 5) {
  map_chr(1:n, ~get_letters(letter_count))
}


# now we can use the same code from 5E, except we use our new alpha roll
# function

dist <- tibble(results = map_chr(rep(1, 1000), roll_alpha_dice_5)) %>% 
  count(results)


ggplot(dist, aes(x = results, y = n)) +
  geom_col() +
  labs(title = "5 Roll Alpha Dice Letter Distribution",
       x = "Letter",
       y = "Letter Count")


```

Clearly, the distribution has shifted more towards the earlier letters rolled per round. This makes sense since choosing the earliest letter from 5 dice (compared to choosing the earliest letter amongst 3 dice) gives a higher likelihood to have an earlier letter rolled and chosen as the winner of each round.

Collaborators: None.
